Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-14 paper.qmd
Removed the spacing

```{r}
#| label: plot-theme
theme_set(theme_linedraw() +
   theme(
     #aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
     panel.background = element_rect(fill = 'transparent', 
                                     colour = NA),
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(), 
     axis.title.x = element_blank(), axis.title.y = element_blank(),
     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
     legend.background = element_rect(fill = 'transparent', 
                                      colour = NA),
     legend.key = element_rect(fill = 'transparent', 
                               colour = NA),
     legend.position = "bottom", 
     legend.title = element_blank(), 
     legend.text = element_text(size=4),
     legend.key.height = unit(0.25, 'cm'),
     legend.key.width = unit(0.25, 'cm')
   )
)
```

@@

nldr1c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr2c <- umap_pbmc |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b", c(0.08, 0.9)) + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("c") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr4c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("d") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr5c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("e") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr6c <- phate_pbmc |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("f") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr7c <- trimap_pbmc |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("g") + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr8c <- pacmap_pbmc |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("h") + 
  theme(plot.margin = margin(0, 0, 0, 0))
```

@@

nldr_two_curvy1 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_two_curvy6 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_two_curvy2 <- umap_two_curvy |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#636363') +
  interior_annotation("b", c(0.08, 0.93)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_two_curvy3 <- phate_two_curvy |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("c") +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_two_curvy4 <- trimap_two_curvy |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("d") +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_two_curvy5 <- pacmap_two_curvy |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1, 
        plot.margin = margin(0, 0, 0, 0))
```

@@

error_plot_two_curvy <-
  plot_mse(error_two_curvy) +
  scale_x_continuous(breaks =
    sort(unique(error_two_curvy$a1))[
      seq(1, length(
        unique(error_two_curvy$a1)), 
        by = 5)]) +
  scale_color_manual(
    values=c('#e41a1c','#ff7f00','#4daf4a', 
             "#a65628",'#636363', '#984ea3')) +
  theme(aspect.ratio = 1.5,
        plot.margin = margin(0, 0, 0, 0))
```

@@

error_plot_pbmc <- plot_mse(error_pbmc) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf')) + 
  theme(plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist1 <- tsne_mnist |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist6 <- tsne_mnist2 |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist2 <- umap_mnist |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b") +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist3 <- phate_mnist |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("c", position = c(0.94, 0.92)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist4 <- trimap_mnist |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("d", position = c(0.86, 0.92)) +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))

@@

nldr_mnist5 <- pacmap_mnist |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0))
```

@@

error_plot_mnist <- plot_mse(error_mnist) +
  scale_x_continuous(breaks = sort(
    unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), 
                                       by = 5), 24)]) +
  scale_color_manual(
    values=c('#a65628', '#984ea3', '#999999', 
             '#377eb8', '#ff7f00', '#e41a1c')) +
  theme(aspect.ratio=1.5,
        plot.margin = margin(0, 0, 0, 0))

@@

top_layout <- wrap_plots(right_top_img, middle_img, right_bottom_img, imge_error_sample_outside, nrow = 4) + theme(plot.margin = margin(0, 0, 0, 0))
full_layout <- highd_error_2d_mnist + top_layout

>>>>

```{r}
#| label: plot-theme
theme_set(theme_linedraw() +
   theme(
     #aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
     panel.background = element_rect(fill = 'transparent', 
                                     colour = NA),
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(), 
     axis.title.x = element_blank(), axis.title.y = element_blank(),
     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
     legend.background = element_rect(fill = 'transparent', 
                                      colour = NA),
     legend.key = element_rect(fill = 'transparent', 
                               colour = NA),
     legend.position = "bottom", 
     legend.title = element_blank(), 
     legend.text = element_text(size=4),
     legend.key.height = unit(0.25, 'cm'),
     legend.key.width = unit(0.25, 'cm'),
     plot.margin = margin(0, 0, 0, 0)
   )
)
```

@@

nldr1c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a")

@@

nldr2c <- umap_pbmc |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b", c(0.08, 0.9))

@@

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("c")

@@

nldr4c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("d")

@@

nldr5c <- tsne_pbmc |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("e")

@@

nldr6c <- phate_pbmc |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("f")

@@

nldr7c <- trimap_pbmc |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("g")

@@

nldr8c <- pacmap_pbmc |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("h") 
```

@@

nldr_two_curvy1 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

nldr_two_curvy6 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

nldr_two_curvy2 <- umap_two_curvy |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#636363') +
  interior_annotation("b", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

nldr_two_curvy3 <- phate_two_curvy |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("c") +
  theme(aspect.ratio = 1)

@@

nldr_two_curvy4 <- trimap_two_curvy |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("d") +
  theme(aspect.ratio = 1)

@@

nldr_two_curvy5 <- pacmap_two_curvy |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1)
```

@@

error_plot_two_curvy <-
  plot_mse(error_two_curvy) +
  scale_x_continuous(breaks =
    sort(unique(error_two_curvy$a1))[
      seq(1, length(
        unique(error_two_curvy$a1)), 
        by = 5)]) +
  scale_color_manual(
    values=c('#e41a1c','#ff7f00','#4daf4a', 
             "#a65628",'#636363', '#984ea3')) +
  theme(aspect.ratio = 1.5)
```

@@

error_plot_pbmc <- plot_mse(error_pbmc) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))

@@

nldr_mnist1 <- tsne_mnist |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist6 <- tsne_mnist2 |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist2 <- umap_mnist |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b") +
  theme(aspect.ratio = 1)

@@

nldr_mnist3 <- phate_mnist |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("c", position = c(0.94, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist4 <- trimap_mnist |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("d", position = c(0.86, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist5 <- pacmap_mnist |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1)
```

@@

error_plot_mnist <- plot_mse(error_mnist) +
  scale_x_continuous(breaks = sort(
    unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), 
                                       by = 5), 24)]) +
  scale_color_manual(
    values=c('#a65628', '#984ea3', '#999999', 
             '#377eb8', '#ff7f00', '#e41a1c')) +
  theme(aspect.ratio=1.5)

@@

top_layout <- wrap_plots(
  right_top_img,
  middle_img,
  right_bottom_img,
  imge_error_sample_outside,
  ncol = 1,
  heights = c(1, 1, 1, 1)
) + plot_layout() + plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))

# Combine with main plot
full_layout <- highd_error_2d_mnist + top_layout +
  plot_layout(widths = c(1, 1)) +
  plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))
